---
title: "Musical Overclaiming: Liking vs. Listening in the Abstract"
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.
:::

# Introduction

The core goal of this analysis is to study "overclaiming" in musical tastes: instances where individuals report that they like a musical genre in the abstract, but that genre is entirely absent when querying the artists and songs actually present on their recent playlists.

To capture this phenomenon, we wrangle survey data to explicitly map abstract genre preferences to concrete listening behavior. We then deploy a multinomial logistic regression framework with two-way clustered standard errors (accounting for both person-level and genre-level variance) to identify which demographic profiles are most prone to overclaiming.

# Data Wrangling and Preparation

The foundational step requires taking raw, self-reported recent streams/artists and mapping them to 20 consistent numeric genre codes. We then align these 20 genres with the 20 corresponding survey questions regarding abstract genre liking.

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(nnet)
library(sandwich)
library(lmtest)
library(marginaleffects)
```

## Binary Indicators for Liking and Listening

We process the core dataset to create two binary indicators for each of the 20 genres:
1. **Liking (`like`)**: Scored as 1 if the respondent rated their liking of the genre as a 5, 6, or 7 on a 7-point Likert scale.
2. **Listening (`listen`)**: Scored as 1 if the genre appeared anywhere in their top 10 reported playlist artists.

We then pivot this data into a long format where each row represents a single person-genre pair.

```{r}
#| label: data-preparation
# Load the base dataset
df_raw <- read_dta("dta/analysis_time_CCC.dta")

# Create IDs
df_raw$id <- 1:nrow(df_raw)

# Map the 20 abstract liking questions to the 20 genre codes
genre_mapping <- list(
  g1=21, g2=22, g3=24, g4=39, g5=26, g6=36, g7=37, g8=23, g9=31, g10=27,
  g11=28, g12=40, g13=29, g14=30, g15=25, g16=32, g17=38, g18=33, g19=34, g20=35
)

# Calculate binary like (5-7) and binary listen indicators
for (g in 1:20) {
  like_col <- paste0("like_music_genres_", genre_mapping[[paste0("g", g)]])
  df_raw[[paste0("like_g", g)]] <- ifelse(df_raw[[like_col]] >= 5 & df_raw[[like_col]] <= 7, 1, 0)
  
  df_raw[[paste0("listen_g", g)]] <- 0
  for (i in 1:10) {
    df_raw[[paste0("listen_g", g)]] <- df_raw[[paste0("listen_g", g)]] | (df_raw[[paste0("genre_final", i)]] == g)
  }
}

# Pivot data to long format for person-genre analysis
# We also create a composite parental education variable
df_long <- df_raw %>%
  mutate(parent_educ = pmax(mom_educ, dad_educ, na.rm = TRUE)) %>%
  select(id, educ2, parent_educ, child_arts, agecat, income, female, urban_rural2, race5, social,
         starts_with("like_g"), starts_with("listen_g")) %>%
  pivot_longer(
    cols = matches("^(like|listen)_g\\d+"),
    names_to = c(".value", "genre_id"),
    names_pattern = "^([a-z]+)_g(\\d+)$"
  ) %>%
  mutate(genre_id = as.factor(genre_id))
```

# Defining the Competing Engagement States

To study overclaiming comprehensively, we frame the behavior as a multi-state competing outcome. For every genre-person pair, a respondent can fall into one of four mutually exclusive engagement states:

1. **Neither**: Does not like it, does not listen to it. (Baseline)
2. **Listen Only (Underclaim)**: Listens to it, but didn't report liking it in the abstract.
3. **Like Only (Overclaim)**: Reports liking it, but it is entirely absent from their listening.
4. **Both (True Engagement)**: Reports liking it AND actively listens to it.

```{r}
#| label: create-outcomes-clean
df_long <- df_long %>%
  mutate(
    engagement_state = case_when(
      like == 0 & listen == 0 ~ "Neither",
      like == 0 & listen == 1 ~ "ListenOnly",
      like == 1 & listen == 0 ~ "LikeOnly",
      like == 1 & listen == 1 ~ "Both"
    ),
    engagement_state = factor(engagement_state, levels = c("Neither", "ListenOnly", "LikeOnly", "Both"))
  )
```

# Descriptive Analysis: Genre-Level Engagement States

Before fitting inferential models, we can visualize the magnitude of these different engagement states across the genres. We calculate the absolute proportion of respondents who fall into each state (Listen Only, Like Only, Both, or Neither) for each of the 20 genres.

```{r}
#| label: descriptive-plot
#| fig-width: 10
#| fig-height: 8
genre_labels <- data.frame(
  genre_id = as.character(1:20),
  genre_name = c(
    "Swing/Big Band", "Bluegrass", "R&B/Blues", "Classic Rock", "Classical",
    "Contemporary Pop", "Contemporary Rock", "Country", "Easy Listening", "Folk",
    "Gospel", "Heavy Metal", "Jazz", "Latin", "Musicals",
    "New Age", "Oldies", "Opera", "Rap/Hip Hop", "Reggae"
  )
)

# Calculate proportions of each engagement state by genre
df_summary <- df_long %>%
  group_by(genre_id, engagement_state) %>%
  summarize(count = n(), .groups = "drop_last") %>%
  mutate(prop = count / sum(count)) %>%
  ungroup() %>%
  left_join(genre_labels, by = "genre_id")

# Order genres by how frequently they are Overclaimed (LikeOnly)
order_by_overclaim <- df_summary %>%
  filter(engagement_state == "LikeOnly") %>%
  arrange(prop) %>%
  pull(genre_name)

df_summary <- df_summary %>%
  mutate(genre_name = factor(genre_name, levels = order_by_overclaim)) %>%
  mutate(engagement_state = factor(engagement_state, 
                                   levels = c("LikeOnly", "ListenOnly", "Both", "Neither"),
                                   labels = c("Like Only (Overclaim)", "Listen Only (Underclaim)", "Both (True Engagement)", "Neither")))

# We will create three separate plots, one for each active engagement state.
# For each plot, the y-axis will be dynamically ordered by the prevalence of that specific state.

df_active <- df_summary %>% filter(engagement_state != "Neither")

plot_state <- function(state_name, title, color) {
  df_filtered <- df_active %>%
    filter(engagement_state == state_name) %>%
    arrange(prop) %>%
    mutate(genre_name = factor(genre_name, levels = genre_name))
  
  avg_prop <- mean(df_filtered$prop)
  
  p <- ggplot(df_filtered, aes(x = prop, y = genre_name)) +
    geom_vline(xintercept = avg_prop, linetype = "dashed", color = "gray40", size = 0.8) +
    geom_segment(aes(x = 0, xend = prop, y = genre_name, yend = genre_name), color = "gray60") +
    geom_point(color = color, size = 4) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme_minimal(base_size = 14) +
    labs(
      title = title,
      subtitle = "Dashed line represents the cross-genre average",
      x = "Percentage of Respondents",
      y = ""
    )
  return(p)
}

# 1. Overclaiming Plot
p_overclaim <- plot_state("Like Only (Overclaim)", "Overclaiming: Like Only", "#E69F00")
ggsave("Plots/Prevalence_Overclaiming.png", p_overclaim, width = 8, height = 6, dpi = 300)

# 2. True Engagement Plot
p_both <- plot_state("Both (True Engagement)", "True Engagement: Like and Listen", "#009E73")
ggsave("Plots/Prevalence_True_Engagement.png", p_both, width = 8, height = 6, dpi = 300)

# 3. Underclaiming Plot
p_underclaim <- plot_state("Listen Only (Underclaim)", "Underclaiming: Listen Only", "#56B4E9")
ggsave("Plots/Prevalence_Underclaiming.png", p_underclaim, width = 8, height = 6, dpi = 300)

# Display the plots in the report
p_overclaim
p_both
p_underclaim
```

The plot above reveals several key findings about genre-level dynamics:
- **Highly Overclaimed**: Easy Listening, Reggae, Classical, and New Age have massive rates of "Like Only" (overclaiming) but almost no "True Engagement" (Both).
- **Highly Consistent**: Contemporary Pop and Contemporary Rock boast the highest rates of "True Engagement" (Both). 
- **Highly Underclaimed**: No genre is dominated by underclaiming, but Contemporary Pop has the highest absolute rate of "Listen Only", meaning many people listen to it without bothering to rate it highly on the survey.

# Defining the Competing Engagement States

# Multinomial Model and Robust Inference

We deploy a fixed-effects multinomial logistic regression (`nnet::multinom`) to model the probability of falling into these states. 

Because our data is in a long format where the same respondent appears 20 times (once for each genre), the observations are not independent. To prevent artificially low $p$-values, we calculate **Two-Way Clustered Standard Errors**. This approach clusters the standard errors simultaneously at the respondent level (`id`) and the genre level (`genre_id`), accounting for variance driven by individual respondent tendencies as well as baseline differences in how heavily specific genres are overclaimed.

```{r}
#| label: multinomial-model
#| results: hide
model_fixed <- multinom(
  engagement_state ~ educ2 + parent_educ + child_arts + agecat + female + income + as.factor(race5),
  data = df_long
)
```

```{r}
#| label: clustered-se
# Calculate two-way clustered standard errors (Person + Genre)
robust_vcov_twoway <- vcovCL(model_fixed, cluster = df_long[, c("id", "genre_id")])
robust_se_twoway <- sqrt(diag(robust_vcov_twoway))

# Manually compute z-scores and p-values
est <- summary(model_fixed)$coefficients
se_mat_twoway <- matrix(robust_se_twoway, nrow = nrow(est), ncol = ncol(est), byrow = TRUE)
z_vals_twoway <- est / se_mat_twoway
p_vals_twoway <- (1 - pnorm(abs(z_vals_twoway), 0, 1)) * 2

# Helper function to extract cleanly and build a nice gt table
library(gt)
library(dplyr)

make_clean_df_twoway <- function(row_idx) {
  df <- data.frame(
    Predictor = colnames(est),
    Estimate = round(est[row_idx, ], 3),
    Robust_SE = round(se_mat_twoway[row_idx, ], 3),
    Z = round(z_vals_twoway[row_idx, ], 2),
    P_Value = round(p_vals_twoway[row_idx, ], 3)
  )
  
  # Clean up predictor names
  df <- df %>%
    mutate(Predictor = case_when(
      Predictor == "(Intercept)" ~ "Intercept",
      Predictor == "educ2" ~ "Education",
      Predictor == "parent_educ" ~ "Parental Education",
      Predictor == "child_arts" ~ "Childhood Arts Exposure",
      Predictor == "agecat" ~ "Age Category",
      Predictor == "female" ~ "Female",
      Predictor == "income" ~ "Household Income",
      Predictor == "as.factor(race5)2" ~ "Race: Black",
      Predictor == "as.factor(race5)3" ~ "Race: Hispanic",
      Predictor == "as.factor(race5)4" ~ "Race: Asian",
      Predictor == "as.factor(race5)5" ~ "Race: Other/Mixed",
      TRUE ~ Predictor
    )) %>%
    mutate(
      Sig = case_when(
        P_Value < 0.001 ~ "***",
        P_Value < 0.01 ~ "**",
        P_Value < 0.05 ~ "*",
        P_Value < 0.1 ~ ".",
        TRUE ~ ""
      )
    )
    
  return(df)
}

format_gt_table <- function(df, title) {
  df %>%
    gt() %>%
    tab_header(
      title = md(title)
    ) %>%
    cols_label(
      Robust_SE = "Robust SE",
      P_Value = "p-value",
      Sig = ""
    ) %>%
    fmt_number(
      columns = c(Estimate, Robust_SE, Z),
      decimals = 3
    ) %>%
    opt_stylize(style = 6, color = "cyan")
}
```

### Predictors of Overclaiming (Like Only)
```{r}
#| label: results-overclaim
#| echo: false
dir.create("Tabs", showWarnings = FALSE)

df_overclaim <- make_clean_df_twoway(2)
gt_overclaim <- format_gt_table(df_overclaim, "**Predictors of Overclaiming (Like Only)**")
gtsave(gt_overclaim, "Tabs/Table_Overclaim.html")
gt_overclaim
```

### Predictors of Underclaiming (Listen Only)
```{r}
#| label: results-underclaim
#| echo: false
df_underclaim <- make_clean_df_twoway(1)
gt_underclaim <- format_gt_table(df_underclaim, "**Predictors of Underclaiming (Listen Only)**")
gtsave(gt_underclaim, "Tabs/Table_Underclaim.html")
gt_underclaim
```

### Predictors of True Engagement (Both)
```{r}
#| label: results-true
#| echo: false
df_true <- make_clean_df_twoway(3)
gt_true <- format_gt_table(df_true, "**Predictors of True Engagement (Like and Listen)**")
gtsave(gt_true, "Tabs/Table_True_Engagement.html")
gt_true
```

The application of two-way clustering reveals a comprehensive picture of genre engagement. **Childhood arts exposure** (`child_arts`) is the strongest and most robust predictor of overclaiming (reporting liking without listening). While it also increases the probability of True Engagement, its effect on Overclaiming is almost twice as strong. Conversely, **Education** (`educ2`) significantly predicts Underclaiming (listening without reporting liking), suggesting that highly educated individuals may consume genres they decline to formally endorse.

# Visualizing the Effect of Arts Exposure

To make these multinomial dynamics intuitive, we calculate the marginal predicted probabilities for each of the four states across the spectrum of childhood arts exposure. We apply our two-way cluster-robust variance-covariance matrix to generate the 95% confidence intervals.

```{r}
#| label: marginal-effects-plot
#| fig-width: 9
#| fig-height: 7
# Generate robust predictions using the two-way clustered vcov matrix
pred_marg <- predictions(
  model_fixed,
  newdata = datagrid(child_arts = 1:7), 
  vcov = robust_vcov_twoway
)

pred_df_rob <- as.data.frame(pred_marg) %>%
  mutate(response.level = case_when(
    group == "Neither" ~ "Neither (No Like, No Listen)",
    group == "ListenOnly" ~ "Listen Only (Underclaim)",
    group == "LikeOnly" ~ "Like Only (Overclaim)",
    group == "Both" ~ "Both (True Engagement)"
  )) %>%
  mutate(response.level = factor(response.level, levels = c(
    "Neither (No Like, No Listen)", 
    "Listen Only (Underclaim)", 
    "Both (True Engagement)", 
    "Like Only (Overclaim)"
  )))

arts_labels <- c("1 (None)", "2", "3", "4 (Moderate)", "5", "6", "7 (Extensive)")

p_margins <- ggplot(pred_df_rob, aes(x = child_arts, y = estimate, color = response.level)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, linewidth = 0.8, alpha = 0.15, position = position_dodge(width = 0.5)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  coord_flip() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = 1:7, labels = arts_labels) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical"
  ) +
  labs(
    title = "Effect of Childhood Arts Exposure on Genre Engagement",
    subtitle = "95% CIs computed via two-way cluster-robust standard errors (Person & Genre)",
    x = "Level of Childhood Arts Exposure",
    y = "Predicted Probability",
    color = "Engagement State"
  )

# Save to Plots directory
dir.create("Plots", showWarnings = FALSE)
ggsave("Plots/ChildArts_Effects_RobustCI_Errorbars.png", p_margins, width = 9, height = 7, dpi = 300)

p_margins
```

As childhood arts exposure increases, respondents genuinely engage with more music (the "True Engagement" estimates increase). However, their propensity to overclaim genres ("Like Only") rises sharply, overtaking the baseline state of ignoring the genre completely. Early arts exposure creates broader actual listening habits, but it creates vastly broader *claimed* listening habits.

### Education and Engagement States

We perform the exact same marginal probability calculation for the highest level of education attained (`educ2`) to see how it affects the competing engagement states.

```{r}
#| label: marginal-effects-educ
#| fig-width: 9
#| fig-height: 7
# Generate robust predictions for education
pred_marg_educ <- predictions(
  model_fixed,
  newdata = datagrid(educ2 = seq(min(df_long$educ2, na.rm=TRUE), max(df_long$educ2, na.rm=TRUE), by=1)), 
  vcov = robust_vcov_twoway
)

pred_df_educ <- as.data.frame(pred_marg_educ) %>%
  mutate(response.level = case_when(
    group == "Neither" ~ "Neither (No Like, No Listen)",
    group == "ListenOnly" ~ "Listen Only (Underclaim)",
    group == "LikeOnly" ~ "Like Only (Overclaim)",
    group == "Both" ~ "Both (True Engagement)"
  )) %>%
  mutate(response.level = factor(response.level, levels = c(
    "Neither (No Like, No Listen)", 
    "Listen Only (Underclaim)", 
    "Both (True Engagement)", 
    "Like Only (Overclaim)"
  )))

educ_labels <- c("Less than HS", "HS Grad", "Some College", "Associates", "BA", "Grad Degree")

p_margins_educ <- ggplot(pred_df_educ, aes(x = educ2, y = estimate, color = response.level)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, linewidth = 0.8, alpha = 0.15, position = position_dodge(width = 0.5)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  coord_flip() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = 1:6, labels = educ_labels) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical"
  ) +
  labs(
    title = "Effect of Education on Genre Engagement",
    subtitle = "95% CIs computed via two-way cluster-robust standard errors (Person & Genre)",
    x = "Education Level",
    y = "Predicted Probability",
    color = "Engagement State"
  )

# Save to Plots directory
ggsave("Plots/Education_Effects_RobustCI_Errorbars.png", p_margins_educ, width = 9, height = 7, dpi = 300)

p_margins_educ
```
