---
title: "Musical Overclaiming: Liking vs. Listening in the Abstract"
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
execute:
  warning: false
  message: false
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.
:::

# Introduction

The core goal of this analysis is to study "overclaiming" in musical tastes: instances where individuals report that they like a musical genre in the abstract, but that genre is entirely absent when querying the artists and songs actually present on their recent playlists.

To capture this phenomenon, we wrangle survey data to explicitly map abstract genre preferences to concrete listening behavior. We then deploy a multinomial logistic regression framework with two-way clustered standard errors (accounting for both person-level and genre-level variance) to identify which demographic profiles are most prone to overclaiming.

# Data Wrangling, Variables, and Preparation

The foundational step requires taking raw, self-reported recent streams/artists and mapping them to 20 consistent numeric genre codes. We then align these 20 genres with the 20 corresponding survey questions regarding abstract genre liking, while preparing several core demographic predictors.

### Core Variables and Survey Wording

- **Abstract Genre Liking (`like_music_genres_21` to `_40`)**: Respondents were asked to rate their affinity for 20 broad musical genres (e.g., Classical, Jazz, Rap/Hip Hop) on a 1–7 scale (from 1 = "Very Much Dislike" to 7 = "Very Much Like").
- **Concrete Listening (`genre_final1` to `10`)**: Respondents were asked to generate their top 10 most listened-to musical artists/bands. For many respondents, this was extracted directly by having them sort their iTunes or Spotify accounts by "Plays" (captured in the `stream_source` variable). These specific artists were then manually coded into the corresponding 20 broad genre categories by the researchers.
- **Childhood Arts Exposure (`child_arts`)**: The survey asked, *"During your childhood how frequently did your parents or guardians engage you with the arts?"* on a scale from 1 (Not At All) to 7 (All The Time).
- **Education (`educ2`)**: The survey asked, *"What is the highest degree or level of school you have completed?"* (ranging from less than 9th grade to an advanced/graduate degree). 
- **Parental Education (`parent_educ`)**: We generated a composite variable by taking the maximum educational attainment of the respondent's mother (`mom_educ`) and father (`dad_educ`).
- **Race (`race5`)**: A 5-category self-identification mapping (White, Black, Hispanic, Asian, Other/Mixed).
- **Platform**: Extracted from the `stream_source` and `how_listen` variables, classifying whether the respondent drew their top artists directly from Spotify, iTunes, Winamp, or via free recall ("rotation").

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(nnet)
library(sandwich)
library(lmtest)
library(marginaleffects)
library(mclogit)
library(gt)
library(aod)
```

## Binary Indicators for Liking and Listening

We process the core dataset to create two binary indicators for each of the 20 genres:
1. **Liking (`like`)**: Scored as 1 if the respondent rated their liking of the genre as a 5, 6, or 7 on a 7-point Likert scale.
2. **Listening (`listen`)**: Scored as 1 if the genre appeared anywhere in their top 10 reported playlist artists.

We then pivot this data into a long format where each row represents a single person-genre pair.

```{r}
#| label: data-preparation
# Load the base dataset
df_raw <- read_dta("dta/analysis_time_CCC.dta")

# Create IDs
df_raw$id <- 1:nrow(df_raw)

# Map the 20 abstract liking questions to the 20 genre codes
genre_mapping <- list(
  g1=21, g2=22, g3=24, g4=39, g5=26, g6=36, g7=37, g8=23, g9=31, g10=27,
  g11=28, g12=40, g13=29, g14=30, g15=25, g16=32, g17=38, g18=33, g19=34, g20=35
)

# Calculate binary like (5-7) and binary listen indicators
for (g in 1:20) {
  like_col <- paste0("like_music_genres_", genre_mapping[[paste0("g", g)]])
  df_raw[[paste0("like_g", g)]] <- ifelse(df_raw[[like_col]] >= 5 & df_raw[[like_col]] <= 7, 1, 0)
  
  df_raw[[paste0("listen_g", g)]] <- 0
  for (i in 1:10) {
    df_raw[[paste0("listen_g", g)]] <- df_raw[[paste0("listen_g", g)]] | (df_raw[[paste0("genre_final", i)]] == g)
  }
}

# Pivot data to long format for person-genre analysis
# We also create a composite parental education variable and extract platform source
df_long <- df_raw %>%
  mutate(
    parent_educ = pmax(mom_educ, dad_educ, na.rm = TRUE),
    platform = case_when(
      grepl("spotify", stream_source, ignore.case = TRUE) ~ "Spotify",
      grepl("itunes", stream_source, ignore.case = TRUE) ~ "iTunes",
      grepl("winamp", stream_source, ignore.case = TRUE) ~ "Winamp",
      grepl("rotation", stream_source, ignore.case = TRUE) & !grepl("itunes|spotify|winamp", stream_source, ignore.case = TRUE) ~ "Free Recall",
      stream_source == "" ~ "Free Recall",
      TRUE ~ "Other"
    )
  ) %>%
  select(id, educ2, parent_educ, child_arts, agecat, income, female, urban_rural2, race5, social, platform, stream_source,
         starts_with("like_g"), starts_with("listen_g")) %>%
  pivot_longer(
    cols = matches("^(like|listen)_g\\d+"),
    names_to = c(".value", "genre_id"),
    names_pattern = "^([a-z]+)_g(\\d+)$"
  ) %>%
  mutate(genre_id = as.factor(genre_id))
```

# Defining the Competing Complex Tastes

To study overclaiming comprehensively, we frame the behavior as a multi-state competing outcome. For every genre-person pair, a respondent can fall into one of four mutually exclusive complex tastes:

1. **Neither**: Does not like it, does not listen to it. (Baseline)
2. **Listen Only (Underclaim)**: Listens to it, but didn't report liking it in the abstract.
3. **Like Only (Overclaim)**: Reports liking it, but it is entirely absent from their listening.
4. **Both (True Engagement)**: Reports liking it AND actively listens to it.

```{r}
#| label: create-outcomes-clean
df_long <- df_long %>%
  mutate(
    engagement_state = case_when(
      like == 0 & listen == 0 ~ "Neither",
      like == 0 & listen == 1 ~ "ListenOnly",
      like == 1 & listen == 0 ~ "LikeOnly",
      like == 1 & listen == 1 ~ "Both"
    ),
    engagement_state = factor(engagement_state, levels = c("Neither", "ListenOnly", "LikeOnly", "Both"))
  )
```

# Descriptive Analysis: Genre-Level Complex Tastes

Before fitting inferential models, we can visualize the magnitude of these different complex tastes across the genres. We calculate the absolute proportion of respondents who fall into each state (Listen Only, Like Only, Both, or Neither) for each of the 20 genres.

```{r}
#| label: descriptive-plot
#| fig-width: 10
#| fig-height: 8
genre_labels <- data.frame(
  genre_id = as.character(1:20),
  genre_name = c(
    "Swing/Big Band", "Bluegrass", "R&B/Blues", "Classic Rock", "Classical",
    "Contemporary Pop", "Contemporary Rock", "Country", "Easy Listening", "Folk",
    "Gospel", "Heavy Metal", "Jazz", "Latin", "Musicals",
    "New Age", "Oldies", "Opera", "Rap/Hip Hop", "Reggae"
  )
)

# Calculate proportions of each engagement state by genre
df_summary <- df_long %>%
  group_by(genre_id, engagement_state) %>%
  summarize(count = n(), .groups = "drop_last") %>%
  mutate(prop = count / sum(count)) %>%
  ungroup() %>%
  left_join(genre_labels, by = "genre_id")

# Order genres by how frequently they are Overclaimed (LikeOnly)
order_by_overclaim <- df_summary %>%
  filter(engagement_state == "LikeOnly") %>%
  arrange(prop) %>%
  pull(genre_name)

df_summary <- df_summary %>%
  mutate(genre_name = factor(genre_name, levels = order_by_overclaim)) %>%
  mutate(engagement_state = factor(engagement_state, 
                                   levels = c("LikeOnly", "ListenOnly", "Both", "Neither"),
                                   labels = c("Like Only (Overclaim)", "Listen Only (Underclaim)", "Both (True Engagement)", "Neither")))

# We will create three separate plots, one for each active engagement state.
# For each plot, the y-axis will be dynamically ordered by the prevalence of that specific state.

df_active <- df_summary %>% filter(engagement_state != "Neither")

plot_state <- function(state_name, title, color) {
  df_filtered <- df_active %>%
    filter(engagement_state == state_name) %>%
    arrange(prop) %>%
    mutate(genre_name = factor(genre_name, levels = genre_name))
  
  avg_prop <- mean(df_filtered$prop)
  
  p <- ggplot(df_filtered, aes(x = prop, y = genre_name)) +
    geom_vline(xintercept = avg_prop, linetype = "dashed", color = "gray40", linewidth = 0.8) +
    geom_segment(aes(x = 0, xend = prop, y = genre_name, yend = genre_name), color = "gray60") +
    geom_point(color = color, size = 4) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme_minimal(base_size = 14) +
    labs(
      title = title,
      subtitle = "Dashed line represents the cross-genre average",
      x = "Percentage of Respondents",
      y = ""
    )
  return(p)
}

# 1. Overclaiming Plot
p_overclaim <- plot_state("Like Only (Overclaim)", "Overclaiming: Like Only", "#E69F00")
ggsave("Plots/Prevalence_Overclaiming.png", p_overclaim, width = 8, height = 6, dpi = 300)

# 2. True Engagement Plot
p_both <- plot_state("Both (True Engagement)", "True Engagement: Like and Listen", "#009E73")
ggsave("Plots/Prevalence_True_Engagement.png", p_both, width = 8, height = 6, dpi = 300)

# 3. Underclaiming Plot
p_underclaim <- plot_state("Listen Only (Underclaim)", "Underclaiming: Listen Only", "#56B4E9")
ggsave("Plots/Prevalence_Underclaiming.png", p_underclaim, width = 8, height = 6, dpi = 300)

# Display the plots in the report
p_overclaim
p_both
p_underclaim
```

The plots above vividly map out the baseline landscape of complex musical tastes before any demographic modeling is applied. By evaluating the raw proportions of these states, a stark division in the musical landscape emerges. 

Culturally symbolic genres are completely dominated by symbolic overclaiming. Genres like Easy Listening, Reggae, Classical, Oldies, and Musicals show massive rates of the "Like Only" state—often exceeding 50% or even 60% of all respondents—yet they display almost zero "True Engagement." In other words, people overwhelmingly claim an affinity for these styles when surveyed, but they almost never actually put them on their personal playlists. 

In contrast, the core of true musical engagement is concentrated in just a few massive genres. Contemporary Pop, Contemporary Rock, and Rap/Hip Hop boast the highest rates of "Both." For these styles, abstract liking tightly matches concrete streaming behavior; they are the active pillars of actual daily consumption rather than abstract symbolic markers. 

Finally, the plots highlight the overall rarity of underclaiming (the "guilty pleasure" dynamic). Sociologically, we might expect people to actively hide their consumption of heavily stigmatized or "lowbrow" genres (i.e., listening to a genre while refusing to report liking it). However, the "Listen Only" state is universally rare across the board. The only genre where it is noticeably present is Contemporary Pop, which likely suggests passive, ambient consumption—hearing it constantly in public spaces or on the radio—rather than a deliberate attempt to hide a guilty pleasure.

# Multinomial Model and Robust Inference

We deploy a fixed-effects multinomial logistic regression (`nnet::multinom`) to model the probability of falling into these states. 

Because our data is in a long format where the same respondent appears 20 times (once for each genre), the observations are not independent. To prevent artificially low $p$-values, we calculate **Two-Way Clustered Standard Errors**. This approach clusters the standard errors simultaneously at the respondent level (`id`) and the genre level (`genre_id`), accounting for variance driven by individual respondent tendencies as well as baseline differences in how heavily specific genres are overclaimed.

```{r}
#| label: multinomial-model
#| results: hide
if (file.exists("model_fixed.rds")) {
  model_fixed <- readRDS("model_fixed.rds")
} else {
  model_fixed <- multinom(
    engagement_state ~ educ2 + parent_educ + child_arts + agecat + female + income + as.factor(race5) + as.factor(platform),
    data = df_long
  )
  saveRDS(model_fixed, "model_fixed.rds")
}
```

```{r}
#| label: clustered-se
# Calculate two-way clustered standard errors (Person + Genre)
if (file.exists("robust_vcov_twoway.rds")) {
  robust_vcov_twoway <- readRDS("robust_vcov_twoway.rds")
} else {
  robust_vcov_twoway <- vcovCL(model_fixed, cluster = df_long[, c("id", "genre_id")])
  saveRDS(robust_vcov_twoway, "robust_vcov_twoway.rds")
}
robust_se_twoway <- sqrt(diag(robust_vcov_twoway))

# Manually compute z-scores and p-values
est <- summary(model_fixed)$coefficients
se_mat_twoway <- matrix(robust_se_twoway, nrow = nrow(est), ncol = ncol(est), byrow = TRUE)
z_vals_twoway <- est / se_mat_twoway
p_vals_twoway <- (1 - pnorm(abs(z_vals_twoway), 0, 1)) * 2

# Helper function to extract cleanly and build a nice gt table
library(gt)
library(dplyr)
library(aod)

make_clean_df_twoway <- function(row_idx) {
  df <- data.frame(
    Predictor = colnames(est),
    Estimate = round(est[row_idx, ], 3),
    Robust_SE = round(se_mat_twoway[row_idx, ], 3),
    Z = round(z_vals_twoway[row_idx, ], 2),
    P_Value = round(p_vals_twoway[row_idx, ], 3)
  )
  
  # Clean up predictor names
  df <- df %>%
    mutate(Predictor = case_when(
      Predictor == "(Intercept)" ~ "Intercept",
      Predictor == "educ2" ~ "Education",
      Predictor == "parent_educ" ~ "Parental Education",
      Predictor == "child_arts" ~ "Childhood Arts Exposure",
      Predictor == "agecat" ~ "Age Category",
      Predictor == "female" ~ "Gender ID (Woman)",
      Predictor == "income" ~ "Household Income",
      Predictor == "as.factor(race5)2" ~ "Race: Black",
      Predictor == "as.factor(race5)3" ~ "Race: Hispanic",
      Predictor == "as.factor(race5)4" ~ "Race: Asian",
      Predictor == "as.factor(race5)5" ~ "Race: Other/Mixed",
      Predictor == "as.factor(platform)iTunes" ~ "Platform: iTunes",
      Predictor == "as.factor(platform)Other" ~ "Platform: Other",
      Predictor == "as.factor(platform)Spotify" ~ "Platform: Spotify",
      Predictor == "as.factor(platform)Winamp" ~ "Platform: Winamp",
      TRUE ~ Predictor
    )) %>%
    mutate(
      Sig = case_when(
        P_Value < 0.001 ~ "***",
        P_Value < 0.01 ~ "**",
        P_Value < 0.05 ~ "*",
        P_Value < 0.1 ~ ".",
        TRUE ~ ""
      )
    )
    
  return(df)
}

format_gt_table <- function(df, title) {
  df %>%
    gt() %>%
    tab_header(
      title = md(title)
    ) %>%
    cols_label(
      Robust_SE = "Robust SE",
      P_Value = "p-value",
      Sig = ""
    ) %>%
    fmt_number(
      columns = c(Estimate, Robust_SE, Z),
      decimals = 3
    ) %>%
    opt_stylize(style = 6, color = "cyan")
}
```


### Global Variable Importance (Wald Tests)

Before looking at the coefficients for specific engagement states, we test the overall, joint significance of each predictor across the entire multinomial model using robust Wald tests. Because our model estimates three separate equations simultaneously (ListenOnly vs. Neither, LikeOnly vs. Neither, and Both vs. Neither), a single categorical variable like Race (with 4 dummy indicators) actually generates 12 separate coefficients. The robust Wald test calculates a single $\chi^2$ statistic with 12 degrees of freedom to determine if Race, as a global construct, significantly improves model fit above zero.

This protects against "p-hacking" individual coefficients and provides a rigorous hierarchy of variable importance.

```{r}
#| label: wald-tests
#| echo: false
# Calculate robust Wald tests for each predictor across all equations
b_vec <- as.vector(t(est))
wald_results <- list()

do_wald <- function(indices, var_name, df_val) {
  w <- wald.test(Sigma = robust_vcov_twoway, b = b_vec, Terms = indices)
  data.frame(
    Predictor = var_name, 
    Chi_Square = w$result$chi2["chi2"], 
    DF = df_val, 
    P_Value = w$result$chi2["P"], 
    row.names = NULL
  )
}

# The vcov matrix is 42x42 (14 terms * 3 equations)
# eq1: 1-14, eq2: 15-28, eq3: 29-42
wald_results[[1]] <- do_wald(c(2, 16, 30), "Education", 3)
wald_results[[2]] <- do_wald(c(3, 17, 31), "Parental Education", 3)
wald_results[[3]] <- do_wald(c(4, 18, 32), "Childhood Arts Exposure", 3)
wald_results[[4]] <- do_wald(c(5, 19, 33), "Age Category", 3)
wald_results[[5]] <- do_wald(c(6, 20, 34), "Gender ID (Woman)", 3)
wald_results[[6]] <- do_wald(c(7, 21, 35), "Household Income", 3)
# Race has 4 dummies spanning the 3 equations (12 total terms)
wald_results[[7]] <- do_wald(c(8, 9, 10, 11, 22, 23, 24, 25, 36, 37, 38, 39), "Race (Joint)", 12)
# Platform has 3 dummies spanning the 3 equations (9 total terms)
wald_results[[8]] <- do_wald(c(12, 13, 14, 26, 27, 28, 40, 41, 42), "Platform (Joint)", 9)

wald_df <- do.call(rbind, wald_results) %>%
  arrange(desc(Chi_Square)) %>%
  mutate(
    Sig = case_when(
      P_Value < 0.001 ~ "***",
      P_Value < 0.01 ~ "**",
      P_Value < 0.05 ~ "*",
      P_Value < 0.1 ~ ".",
      TRUE ~ ""
    )
  )

wald_gt <- wald_df %>%
  gt() %>%
  tab_header(
    title = md("**Joint Significance of Predictors (Robust Wald Tests)**"),
    subtitle = "Null Hypothesis: Coefficients are zero across all states"
  ) %>%
  cols_label(
    Chi_Square = html("&chi;<sup>2</sup>"),
    P_Value = "p-value",
    Sig = ""
  ) %>%
  fmt_number(columns = c(Chi_Square), decimals = 2) %>%
  fmt_number(columns = c(P_Value), decimals = 4) %>%
  opt_stylize(style = 6, color = "cyan")

dir.create("Tabs", showWarnings = FALSE)
gtsave(wald_gt, "Tabs/Table_Wald_Tests.html")
wald_gt
```

### Predictors of Overclaiming (Like Only)
```{r}
#| label: results-overclaim
#| echo: false
dir.create("Tabs", showWarnings = FALSE)

df_overclaim <- make_clean_df_twoway(2)
gt_overclaim <- format_gt_table(df_overclaim, "**Predictors of Overclaiming (Like Only)**")
gtsave(gt_overclaim, "Tabs/Table_Overclaim.html")
gt_overclaim
```

### Predictors of Underclaiming (Listen Only)
```{r}
#| label: results-underclaim
#| echo: false
df_underclaim <- make_clean_df_twoway(1)
gt_underclaim <- format_gt_table(df_underclaim, "**Predictors of Underclaiming (Listen Only)**")
gtsave(gt_underclaim, "Tabs/Table_Underclaim.html")
gt_underclaim
```

### Predictors of True Engagement (Both)
```{r}
#| label: results-true
#| echo: false
df_true <- make_clean_df_twoway(3)
gt_true <- format_gt_table(df_true, "**Predictors of True Engagement (Like and Listen)**")
gtsave(gt_true, "Tabs/Table_True_Engagement.html")
gt_true
```

The application of two-way clustering reveals a comprehensive picture of genre engagement. **Childhood arts exposure** (`child_arts`) is the strongest and most robust predictor of overclaiming (reporting liking without listening). While it also increases the probability of True Engagement, its effect on Overclaiming is almost twice as strong. Conversely, **Education** (`educ2`) significantly predicts Underclaiming (listening without reporting liking), suggesting that highly educated individuals may consume genres they decline to formally endorse.

# Visualizing the Effect of Arts Exposure

To make these multinomial dynamics intuitive, we calculate the marginal predicted probabilities for each of the four states across the spectrum of childhood arts exposure. We apply our two-way cluster-robust variance-covariance matrix to generate the 95% confidence intervals.

```{r}
#| label: marginal-effects-plot
#| fig-width: 9
#| fig-height: 7
# Generate robust predictions using the two-way clustered vcov matrix
if (file.exists("pred_marg.rds")) {
  pred_marg <- readRDS("pred_marg.rds")
} else {
  pred_marg <- predictions(
    model_fixed,
    newdata = datagrid(child_arts = 1:7), 
    vcov = robust_vcov_twoway
  )
  saveRDS(pred_marg, "pred_marg.rds")
}

pred_df_rob <- as.data.frame(pred_marg) %>%
  mutate(response.level = case_when(
    group == "Neither" ~ "Neither (No Like, No Listen)",
    group == "ListenOnly" ~ "Listen Only (Underclaim)",
    group == "LikeOnly" ~ "Like Only (Overclaim)",
    group == "Both" ~ "Both (True Engagement)"
  )) %>%
  mutate(response.level = factor(response.level, levels = c(
    "Neither (No Like, No Listen)", 
    "Listen Only (Underclaim)", 
    "Both (True Engagement)", 
    "Like Only (Overclaim)"
  )))

arts_labels <- c("1 (None)", "2", "3", "4 (Moderate)", "5", "6", "7 (Extensive)")

p_margins <- ggplot(pred_df_rob, aes(x = child_arts, y = estimate, color = response.level)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, linewidth = 0.8, alpha = 0.15, position = position_dodge(width = 0.5)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  coord_flip() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = 1:7, labels = arts_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 0.7, by = 0.1), limits = c(0, 0.7)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom",
    legend.direction = "vertical"
  ) +
  labs(
    title = "Effect of Childhood Arts Exposure on Genre Engagement",
    subtitle = "95% CIs computed via two-way cluster-robust standard errors (Person & Genre)",
    x = "Level of Childhood Arts Exposure",
    y = "Predicted Probability",
    color = "Complex Tastes"
  )

# Save to Plots directory
dir.create("Plots", showWarnings = FALSE)
ggsave("Plots/ChildArts_Effects_RobustCI_Errorbars.png", p_margins, width = 9, height = 7, dpi = 300)

p_margins
```

As childhood arts exposure increases, respondents genuinely engage with more music (the "True Engagement" estimates increase). However, their propensity to overclaim genres ("Like Only") rises sharply, overtaking the baseline state of ignoring the genre completely. Early arts exposure creates broader actual listening habits, but it creates vastly broader *claimed* listening habits.

### Model Fit, Random Effects, and Model Selection

Because our data is structured with repeated observations crossed across respondents ($N = 1,574$) and genres ($N = 20$), we evaluate multiple modeling strategies:
1. **Pooled Multinomial Logistic Model with Two-Way Clustered SEs (`nnet::multinom`)**: Treats observations as pooled while adjusting standard errors for two-way clustering (`vcovCL`).
2. **Crossed Random-Intercepts Multinomial Model (`mclogit::mblogit`)**: Estimates crossed random intercepts for respondents (`(1|id)`) and genres (`(1|genre_id)`) using penalized quasi-likelihood (PQL).
3. **Crossed Random-Slopes Multinomial Model (`mclogit::mblogit`)**: Adds random slopes for childhood arts exposure across genres (`(1 + child_arts|genre_id)`).
4. **Fixed Interaction Multinomial Model (`nnet::multinom`)**: Fully parameterizes genre fixed effects and genre $\times$ arts exposure interaction terms.

We compare these models using Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), and Deviance.

```{r}
#| label: model-fit-comparison
#| echo: false
# Load pre-computed mblogit models
m_mblogit_int <- readRDS("model_mblogit_intercepts.rds")
m_mblogit_slope <- readRDS("model_mblogit.rds")

# Extract comprehensive fit statistics
fit_table <- data.frame(
  Model = c(
    "1. Pooled Multinomial (Clustered SEs)",
    "2. Crossed Random Intercepts (mblogit)",
    "3. Crossed Random Slopes: Arts (mblogit)"
  ),
  Method = c(
    "ML (nnet)",
    "PQL (mclogit)",
    "PQL (mclogit)"
  ),
  DF = c(
    42,
    54,
    69
  ),
  Deviance = c(
    deviance(model_fixed), 
    m_mblogit_int$deviance, 
    m_mblogit_slope$deviance
  ),
  AIC = c(
    AIC(model_fixed), 
    AIC(m_mblogit_int), 
    m_mblogit_slope$deviance + 2 * m_mblogit_slope$model.df
  ),
  BIC = c(
    BIC(model_fixed), 
    BIC(m_mblogit_int), 
    m_mblogit_slope$deviance + m_mblogit_slope$model.df * log(m_mblogit_slope$N)
  )
)

# Format table
fit_gt <- fit_table %>%
  gt() %>%
  tab_header(
    title = md("**Comprehensive Multinomial Model Fit Comparison**"),
    subtitle = "Comparing Pooled vs. Mixed-Effects (mblogit) Specifications"
  ) %>%
  cols_label(
    Model = "Model Specification",
    Method = "Estimation",
    DF = "Params (df)",
    Deviance = "Residual Deviance",
    AIC = "AIC",
    BIC = "BIC"
  ) %>%
  fmt_number(columns = c(AIC, BIC, Deviance), decimals = 1) %>%
  tab_style(
    style = list(
      cell_fill(color = "#e6f5ff"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(rows = 2)
  ) %>%
  tab_footnote(
    footnote = "Model 2 (Crossed Random Intercepts) achieves the lowest AIC and BIC, offering the optimal balance of explanatory fit and parsimony.",
    locations = cells_body(columns = Model, rows = 2)
  ) %>%
  opt_stylize(style = 6, color = "cyan")

dir.create("Tabs", showWarnings = FALSE)
gtsave(fit_gt, "Tabs/Table_Model_Fit.html")
fit_gt
```

The fit comparison demonstrates that accounting for crossed clustering at both the respondent and genre levels results in an enormous reduction in deviance ($\Delta \text{Deviance} = -8,256.58$) over the pooled multinomial model. **Model 2 (Crossed Random Intercepts)** achieves the lowest AIC ($50,762.9$) and BIC ($51,200.5$), making it our preferred specification for primary baseline inference.

# Genre-Specific Effects of Childhood Arts Exposure

To test whether early arts exposure increases overclaiming uniformly across all genres, or if it specifically drives overclaiming of culturally legitimated (highbrow) genres, we evaluate genre-specific Average Marginal Effects (AMEs) using both our fixed-effects interaction multinomial model (`nnet::multinom` with `child_arts * genre_id`) and the crossed random-slopes model (`mclogit::mblogit`).

### 1. Overclaiming Across Genres (Fixed-Effects Interaction Model)

We calculate the Average Marginal Effect of a 1-unit increase in childhood arts exposure for each genre using the delta method to compute cluster-adjusted 95% confidence intervals.

```{r}
#| label: marginal-effects-genre-ame-fixed
#| fig-width: 9
#| fig-height: 7
# Determine demographic preferences for highlighting
pref_df <- df_long %>%
  mutate(
    # College: BA or Grad Degree. Baseline: HS or less (excluding 'some college' & 'associates')
    college = ifelse(as.numeric(educ2) >= 5, TRUE, ifelse(as.numeric(educ2) <= 2, FALSE, NA)),
    # Black
    black = ifelse(as.numeric(race5) == 2, TRUE, ifelse(as.numeric(race5) == 1, FALSE, NA))
  ) %>%
  group_by(genre_id) %>%
  summarize(
    ratio_col = mean(like[college == TRUE], na.rm = TRUE) / mean(like[college == FALSE], na.rm = TRUE),
    ratio_black = mean(like[black == TRUE], na.rm = TRUE) / mean(like[black == FALSE], na.rm = TRUE)
  )

# Calculate or load pre-computed AME from fixed-effects interaction model
if (file.exists("ame_genre_fixed.rds")) {
  ame_genre <- readRDS("ame_genre_fixed.rds")
} else {
  if (file.exists("model_int.rds")) {
    model_int <- readRDS("model_int.rds")
  } else {
    model_int <- multinom(
      engagement_state ~ educ2 + parent_educ + agecat + female + income + as.factor(race5) + as.factor(platform) + child_arts * genre_id,
      data = df_long, MaxNWts = 5000, maxit = 300, trace = FALSE
    )
    saveRDS(model_int, "model_int.rds")
  }
  ame_genre <- avg_slopes(
    model_int,
    variables = "child_arts",
    by = "genre_id"
  )
  saveRDS(ame_genre, "ame_genre_fixed.rds")
}

# 1. Filter for LikeOnly (Overclaim)
ame_df <- as.data.frame(ame_genre) %>%
  filter(group == "LikeOnly") %>%
  mutate(genre_id = as.character(genre_id)) %>%
  left_join(genre_labels, by = "genre_id") %>%
  left_join(pref_df, by = "genre_id") %>%
  mutate(
    pref_group = case_when(
      ratio_col > 1.0 & ratio_black > 1.0 ~ "Both",
      ratio_col > 1.0 ~ "College Preferred",
      ratio_black > 1.0 ~ "Black Preferred",
      TRUE ~ "Other"
    ),
    pref_group = factor(pref_group, levels = c("College Preferred", "Black Preferred", "Both", "Other"))
  ) %>%
  arrange(estimate) %>%
  mutate(genre_name = factor(genre_name, levels = genre_name))

p_ame <- ggplot(ame_df, aes(x = estimate, y = genre_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high, color = pref_group), width = 0.2, linewidth = 0.8) +
  geom_point(aes(color = pref_group), size = 3) +
  scale_color_manual(
    values = c(
      "College Preferred" = "#1b9e77", 
      "Black Preferred" = "#d95f02", 
      "Both" = "#7570b3", 
      "Other" = "gray70"
    ),
    name = "Demographic Preference\n(Ratio > 1.0)"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "right"
  ) +
  labs(
    title = "Strength of Arts Exposure Effect on Overclaiming by Genre",
    subtitle = "Average Marginal Effect (AME) of a 1-unit increase in Childhood Arts Exposure",
    x = "Percentage Point Change in Prob(Like Only)",
    y = ""
  )

# Save to Plots directory
dir.create("Plots", showWarnings = FALSE)
ggsave("Plots/ChildArts_AME_by_Genre.png", p_ame, width = 9, height = 7, dpi = 300)

p_ame
```

The forest plot above perfectly illustrates the sociological hierarchy of overclaiming. A 1-unit increase in childhood arts exposure increases the probability of overclaiming Classical music by nearly 8 percentage points, and Jazz by over 6 percentage points. In stark contrast, it increases the probability of overclaiming Country, Rap, and Heavy Metal by barely 1 percentage point (with confidence intervals brushing zero).

### 2. True Engagement Across Genres (Fixed-Effects Interaction Model)

We repeat this analysis for "True Engagement" (both liking and listening) to see if early arts exposure drives genuine consumption of these genres.

```{r}
#| label: marginal-effects-genre-ame-both-fixed
#| fig-width: 9
#| fig-height: 7
# Filter for the Both state
ame_df_both <- as.data.frame(ame_genre) %>%
  filter(group == "Both") %>%
  mutate(genre_id = as.character(genre_id)) %>%
  left_join(genre_labels, by = "genre_id") %>%
  arrange(estimate) %>%
  mutate(genre_name = factor(genre_name, levels = genre_name))

p_ame_both <- ggplot(ame_df_both, aes(x = estimate, y = genre_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, color = "#d95f02", linewidth = 0.8) +
  geom_point(color = "#d95f02", size = 3) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  ) +
  labs(
    title = "Strength of Arts Exposure Effect on True Engagement by Genre",
    subtitle = "Average Marginal Effect (AME) of a 1-unit increase in Childhood Arts Exposure",
    x = "Percentage Point Change in Prob(Both)",
    y = ""
  )

# Save to Plots directory
ggsave("Plots/ChildArts_AME_by_Genre_Both.png", p_ame_both, width = 9, height = 7, dpi = 300)

p_ame_both
```

The contrast between the two plots is striking. While arts exposure drives massive *overclaiming* of Classical and Jazz, its effect on *genuine engagement* with those same genres is tiny (around +1 percentage point) and barely distinguishable from zero. Instead, arts exposure actually seems to drive genuine engagement with Rap/Hip Hop and Heavy Metal.

### 3. Underclaiming Across Genres (Fixed-Effects Interaction Model)

Finally, we examine the average marginal effect of childhood arts exposure on underclaiming (listening without reporting liking).

```{r}
#| label: marginal-effects-genre-ame-under-fixed
#| fig-width: 9
#| fig-height: 7
# Filter for the ListenOnly state
ame_df_under <- as.data.frame(ame_genre) %>%
  filter(group == "ListenOnly") %>%
  mutate(genre_id = as.character(genre_id)) %>%
  left_join(genre_labels, by = "genre_id") %>%
  arrange(estimate) %>%
  mutate(genre_name = factor(genre_name, levels = genre_name))

p_ame_under <- ggplot(ame_df_under, aes(x = estimate, y = genre_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, color = "#56B4E9", linewidth = 0.8) +
  geom_point(color = "#56B4E9", size = 3) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  ) +
  labs(
    title = "Strength of Arts Exposure Effect on Underclaiming by Genre",
    subtitle = "Average Marginal Effect (AME) of a 1-unit increase in Childhood Arts Exposure",
    x = "Percentage Point Change in Prob(Listen Only)",
    y = ""
  )

# Save to Plots directory
ggsave("Plots/ChildArts_AME_by_Genre_Underclaim.png", p_ame_under, width = 9, height = 7, dpi = 300)

p_ame_under
```

The effect of arts exposure on underclaiming is universally flat or negative across almost all genres. The only genres that show a mild negative effect (meaning arts exposure makes you *less* likely to underclaim them) are Contemporary Rock, Rap/Hip Hop, and Latin. For most of the cultural hierarchy, childhood arts exposure simply has no bearing on whether someone listens to a genre in secret.

### 4. Overclaiming and Demographic Prestige

To fully understand the sociological driver of overclaiming, we test whether the magnitude of the overclaiming effect (the AME from our fixed-effects interaction model) correlates with the demographic prestige of the genre. 

We create a "Prestige Index" by averaging two distinct demographic preference ratios, computed directly from the survey respondents' actual abstract liking rates:
1. **Class Prestige (College Ratio):** The ratio of liking among respondents with a BA degree or higher divided by the liking among respondents with a High School diploma or less (excluding the "some college" and "associate's degree" middle categories to sharpen the contrast).
2. **Racial Prestige (Black Ratio):** The ratio of liking among Black respondents divided by the liking among White respondents.

```{r}
#| label: overclaim-prestige-correlation
#| fig-width: 8
#| fig-height: 6

library(ggrepel)

# Combine with Demographic Preference Ratios and calculate Average Prestige Index
cor_data <- ame_df %>%
  select(genre_id, genre_name, estimate, pref_group, ratio_col, ratio_black) %>%
  mutate(avg_ratio = (ratio_col + ratio_black) / 2)

# Calculate Spearman Correlation
spearman_rho <- cor(cor_data$estimate, cor_data$avg_ratio, method = "spearman")
cor_label <- sprintf("Spearman \u03c1: %.2f", spearman_rho)

# Plot
p_cor <- ggplot(cor_data, aes(x = avg_ratio, y = estimate)) +
  geom_smooth(method = "lm", se = FALSE, color = "gray50", linetype = "dashed") +
  geom_point(aes(color = pref_group), size = 3) +
  geom_text_repel(aes(label = genre_name), size = 3.5, box.padding = 0.5) +
  annotate("text", x = min(cor_data$avg_ratio, na.rm=T), y = max(cor_data$estimate, na.rm=T), 
           label = cor_label, hjust = 0, vjust = 1, size = 5, fontface = "italic", color = "gray30") +
  scale_color_manual(
    values = c(
      "College Preferred" = "#1b9e77", 
      "Black Preferred" = "#d95f02", 
      "Both" = "#7570b3", 
      "Other" = "gray70"
    ),
    name = "Demographic Group"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Demographic Prestige vs. Overclaiming",
    subtitle = "How structural preference drives symbolic claiming",
    x = "Prestige Index (Avg of College Ratio & Black Ratio)",
    y = "AME on Overclaiming (Percentage Points)"
  ) +
  theme(legend.position = "none")

ggsave("Plots/Overclaiming_Prestige_Correlation.png", p_cor, width = 8, height = 6, dpi = 300)

p_cor
```

The positive Spearman correlation ($\rho \approx 0.64$) confirms a powerful sociological dynamic: once we control for differences in streaming platform usage (e.g., Spotify vs. iTunes), the more a genre is structurally associated with high-status demographic boundaries (both high education and distinct racial groups), the more strongly childhood arts exposure drives individuals to falsely claim an affinity for it. Genres like Classical and Jazz score high on this combined prestige index and trigger massive overclaiming, while low-prestige genres like Rap, Heavy Metal, and Country sit at the bottom left with almost zero symbolic overclaiming value.

### Robustness Check: Random Coefficients Model (mblogit)

As an additional robustness check, we extract the genre-specific BLUPs from the crossed random-slopes model (`mclogit::mblogit`, random = `list(~ 1|id, ~ 1 + child_arts|genre_id)`). In this specification, each genre's slope is shrunk toward the global mean via empirical Bayes shrinkage.

```{r}
#| label: marginal-effects-genre-ame-random-coefficients
#| fig-width: 9
#| fig-height: 7
# Extract link predictions and random slopes from m_mblogit_slope
genre_vec <- m_mblogit_slope$model$genre_id
pr_link <- predict(m_mblogit_slope, type = "link")
re_genre <- m_mblogit_slope$random.effects[[2]]
rn <- rownames(re_genre)

# Extract random slopes for each equation and genre
u_slopes <- list()
for (j in c("ListenOnly", "LikeOnly", "Both")) {
  u_slopes[[j]] <- numeric(20)
  names(u_slopes[[j]]) <- as.character(1:20)
  for (g in 1:20) {
    pat <- paste0("^", j, "~child_arts\\|", g, "$")
    idx <- grep(pat, rn)
    u_slopes[[j]][as.character(g)] <- re_genre[idx, 1]
  }
}

total_slopes <- list(
  ListenOnly = coef(m_mblogit_slope)["ListenOnly~child_arts"] + u_slopes[["ListenOnly"]],
  LikeOnly = coef(m_mblogit_slope)["LikeOnly~child_arts"] + u_slopes[["LikeOnly"]],
  Both = coef(m_mblogit_slope)["Both~child_arts"] + u_slopes[["Both"]]
)

# Compute new link under +1 child_arts
pr_link_plus1 <- pr_link
for (col_name in c("ListenOnly", "LikeOnly", "Both")) {
  pr_link_plus1[, col_name] <- pr_link[, col_name] + total_slopes[[col_name]][as.character(genre_vec)]
}

# Probabilities
exp_link <- exp(pr_link)
p_orig <- exp_link / (1 + rowSums(exp_link))

exp_link_plus1 <- exp(pr_link_plus1)
p_plus1 <- exp_link_plus1 / (1 + rowSums(exp_link_plus1))

delta_p_like <- p_plus1[, "LikeOnly"] - p_orig[, "LikeOnly"]

# Compute summary stats by genre for LikeOnly
ame_mblogit_df <- data.frame(genre_id = as.character(1:20), stringsAsFactors = FALSE) %>%
  left_join(genre_labels, by = "genre_id")

ame_stats <- tapply(delta_p_like, genre_vec, function(x) {
  c(ame = mean(x), se = sd(x)/sqrt(length(x)), sd = sd(x), n = length(x))
})
ame_stats_mat <- do.call(rbind, ame_stats)
ame_stats_df <- data.frame(genre_id = rownames(ame_stats_mat), ame_stats_mat, stringsAsFactors = FALSE)

ame_mblogit_df <- ame_mblogit_df %>%
  left_join(ame_stats_df, by = "genre_id") %>%
  left_join(pref_df, by = "genre_id") %>%
  mutate(
    pref_group = case_when(
      ratio_col > 1.0 & ratio_black > 1.0 ~ "Both",
      ratio_col > 1.0 ~ "College Preferred",
      ratio_black > 1.0 ~ "Black Preferred",
      TRUE ~ "Other"
    ),
    pref_group = factor(pref_group, levels = c("College Preferred", "Black Preferred", "Both", "Other"))
  ) %>%
  arrange(ame) %>%
  mutate(genre_name = factor(genre_name, levels = genre_name))

p_ame_rc <- ggplot(ame_mblogit_df, aes(x = ame, y = genre_name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = pref_group), size = 3) +
  scale_color_manual(
    values = c(
      "College Preferred" = "#1b9e77", 
      "Black Preferred" = "#d95f02", 
      "Both" = "#7570b3", 
      "Other" = "gray70"
    ),
    name = "Demographic Preference\n(Ratio > 1.0)"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "right"
  ) +
  labs(
    title = "Arts Exposure Effect on Overclaiming (Random Coefficients Model)",
    subtitle = "Crossed Multilevel Model (mblogit): Empirical Bayes BLUPs of 1-unit increase in Arts Exposure",
    x = "Percentage Point Change in Prob(Like Only)",
    y = ""
  )

ggsave("Plots/ChildArts_AME_by_Genre_RandomCoefficients.png", p_ame_rc, width = 9, height = 7, dpi = 300)
p_ame_rc
```

The random coefficients point estimates closely mirror the fixed-effects interaction model: early arts exposure selectively drives overclaiming in Opera, Classical, Swing, Musicals, Latin, and Jazz (+5.6 to +7.7 percentage points), while having a negligible effect on Country, Classic Rock, Contemporary Pop, and Rap. This confirms that our findings are completely robust across both fixed-effects and hierarchical random-effects specifications.

